Symmetric Circuits for Rank Gates

### 2018-02-13 sections/translating-formulas-to-cicuits.tex
FEB 13

\subsection{Translating FPR to Symmetric Matrix Circuits}
In this subsection we translate $\FPR$-formulas into families of transparent
rank-circuits. In particular, we now show that for $\theta(\vec{x}) \in
\FPR[\tau]$ there exists a $P$-uniform family of symmetric rank-circuits
$(C_n)_{n \in \nats}$ such that the query computed by $(C_n)_{n \in \nats}$ is
equal to the query defined by $\theta (\vec{x})$.

\begin{thm}
  For any relational vocabulary $\tau$ and each $\theta(\vec{x}) \in \FPR[\tau]$
  there is a map $n \mapsto C_n$, where $C_n$ is a transparent symmetric
  rank-circuit, and is computable in time polynomial in $n$. The family of
  circuits $(C_n)_{n \in \nats}$ defined by this mapping computes the same query
  as the one defined by $\theta(\vec{x})$.
  \label{thm:translating-formulas-to-circuits}
\end{thm}

\begin{proof}
  We say that a circuit $C$ \emph{translates} a formula $\psi(\vec{x})$ for $n
  \in \nats$ if $C$ is a transparent symmetric rank-circuit of order $n$ and $C$
  computes the query defined by $\psi(\vec{x})$ on structures of size $n$. We
  say a family of circuits $(C_n)_{n \in \nats}$ \emph{translates} a formula
  $\psi (\vec{x})$ if $(C_n)_{n \in \nats}$ is $\PT$-uniform and for all $n \in
  \nats$, $C_n$ translates $\psi(\vec{x})$ for $n$.

  We have from Anderson and Dawar~\cite{AndersonD17} that for every formula
  $\phi(\vec{x}) \in \FPC[\tau]$ there is a $\PT$-uniform family of symmetric
  circuits $(C^{\phi}_n)_{n \in \nats}$ with unique labels and symmetric gates
  that translates $\phi(\vec{x})$.

  Let $\theta(\vec{x}) \in \FPR[\tau]$ and $t := \width (\theta)$. We begin by
  showing that there exists a $P$-uniform family of formulas
  $(\theta(\vec{x}))_{n \in \nats}$ such that for each $n \in \nats$, we have
  $\theta_n(\vec{x}) \in \FOrk[\tau]$, $\width(\theta_n) \leq 4 t$, and
  $\theta_n(\vec{x})$ is equivalent to $\theta(\vec{x})$ on structures of size
  $n$. This result follows from two observations.

>>>>

\begin{definition}
  The function $\op{tag-inputs}$ maps a circuit $C := \langle G, \Omega, \Sigma,
  \Lambda, L \rangle$, a relation symbol $R \in \tau$, and $H \subseteq G$
\end{definition}

\subsection{Translating FPR to Symmetric Matrix Circuits}
% In this subsection we translate $\FPR$-formulas into families of transparent
% rank-circuits. In particular, we now show that for $\theta(\vec{x}) \in
% \FPR[\tau]$ there exists a $P$-uniform family of symmetric rank-circuits
% $(C_n)_{n \in \nats}$ such that the query computed by $(C_n)_{n \in \nats}$ is
% equal to the query defined by $\theta (\vec{x})$.

We say that a circuit $C$ \emph{translates} a formula $\psi(\vec{x})$ for $n \in
\nats$ if $C$ is a transparent symmetric rank-circuit of order $n$, $C$ is
defined over the same vocabulary as $\psi$, and $C$ computes the query defined
by $\psi(\vec{x})$ on structures of size $n$. We say a family of circuits
$(C_n)_{n \in \nats}$ \emph{translates} a formula $\psi (\vec{x})$ if $(C_n)_{n
  \in \nats}$ is $\PT$-uniform and for all $n \in \nats$, $C_n$ translates
$\psi(\vec{x})$ for $n$.

We now prove that for every $\FOrk$-formula $\theta(\vec{x})$ and $n \in \nats$
there exists a circuit $C$ that translates $\theta(\vec{x})$ for $n$. We
construct $C$ by including for each sub-formula $\psi \in \cl{\theta}$ and
assignment $\alpha \in [n]^{\free{\psi}}$ a gate $g_{\psi[\alpha]}$. The
children of $g_{\psi[\alpha]}$ are the gates of the form $g_{\phi[\beta]}$,
where $\phi$ is an immediate sub-formula of $\psi$ and $\beta$ is an assignment
to the free variables of $\psi$ that is compatible with the assignment $\alpha$.
The gate $g_{\psi[\alpha]}$ for an input structure over the universe $[n]$
evaluates to one if, and only if, $\psi$ holds for the input structure with the
assignment $\alpha$. We now construct the translation of $\theta$ for $n$, and
prove that the construction can be completed in a time bound that depends on $n$
and certain properties of the formula $\theta$.

\begin{lem}
  There is an algorithm that takes in a $\FOrk$-formula $\theta(\vec{x})$ and $n
  \in \nats$ and outputs a rank-circuit $C$ that translates $\theta(\vec{x})$
  for $n$. Moreover, there is a polynomial in $p$ such that for an input
  $(\theta(\vec{x}), n)$ the algorithm terminates in at most
  $p(n^{\width(\theta)}\vert \cl{\theta} \vert)$ many steps.
  \label{lem:translating-FOrk}
\end{lem}
\begin{proof}
  Let $\tau$ be the vocabulary of $\theta(\vec{x})$. Let $k := \width(\theta)$.
  Suppose $\tau$ is empty. Since the vocabulary of $\tau$ is empty, all of the
  atomic formula in $\theta(\vec{x})$ are of the form $y_1 = y_2$, where $y_1$
  and $y_2$ are variables. It can be shown that, for an assignment $\vec{a}$ to
  $\vec{x}$, we can evaluate $[n] \models \theta[\vec{a}]$ in time polynomial in
  $n^{k} \vert \cl{\theta} \vert$. Let $f : [n]^k \rightarrow \{0,1\}$ be
  defined such that $f (\vec{a}) = 1$ if, and only if, $[n] \models
  \theta[\vec{a}]$. Let $C := \langle G, \Sigma, \Omega, \Lambda, L \rangle$,
  where
  \begin{itemizens}
  \item $G = \{g_0, g_1\} \cup \{g_{\vec{a}} : \vec{a} \in [n]^{k}\}$,
  \item $\Sigma (g_0) = 0$, $\Sigma (g_1) = 1$ and $\Sigma (g_{\vec{a}}) =
    \land[1]$,
  \item $\Omega : [n]^k \rightarrow G$ is defined by $\Omega(\vec{a}) =
    g_{\vec{a}}$ for all $\vec{a} \in [n]^k$,
  \item $\Lambda = \emptyset$, and
  \item for $\vec{a} \in [n]^k$, $L(g_{\vec{a}}) : [1] \rightarrow G$ is defined
    by $L(g_{\vec{a}})(1) = g_{f(\vec{a})}$.
  \end{itemizens}

  Since $C$ has symmetric gates, $C$ is transparent. Notice that, since
  $\theta(\vec{x})$ defines a query on structures over the empty vocabulary, we
  have for all $\vec{a} \in [n]^k$, $\sigma \in \sym_n$, $f(\sigma \vec{a}) =
  f(\vec{a})$. For $\sigma \in \sym_n$, let $\pi_\sigma : G \rightarrow G$ be a
  bijection that fixes the constant gates and such that $\pi_{\sigma}
  g_{\vec{a}} = g_{\sigma \vec{a}}$ for all $\vec{a} \in [n]^k$. Then
  $\pi_\sigma \Omega (\vec{a}) = \pi_\sigma g_{\vec{a}} = g_{\sigma \vec{a}} =
  \Omega (\sigma \vec{a})$. It follows that $\pi_\sigma$ is an automorphism of
  the circuit extending $\sigma$, and $C$ is symmetric. We thus have that $C$
  translates $\theta(\vec{x})$ for $n$.

  Suppose $\tau$ is not empty. Let $R \in \tau$. We first define two helper
  formulas. For a variable $y$ let $\op{no-op}(y) := (R(y, y) \land (\neg R(y,
  y)))$. For a sequence of variables $\vec{y} = (y_1, \ldots, y_m)$ let
  $\op{tag} (\vec{y}) := \op{no-op}(y_1) \land (\op{no-op}(y_2) \land (
  \op{no-op}(y_2) \land ( \ldots \land (\op{no-op}(y_m)) \ldots)))$. Let
  $\lambda (\vec{x})$ be constructed from $\theta(\vec{x})$ by replacing each
  sub-formula $\psi(\vec{y})$ of the form $\rank^r_p \vec{w} \vec{z} . \phi$
  with the formula $\rank^r_p \vec{w}\vec{z} . ((\forall u . u = u) \land \phi)
  \land \op{tag}(\vec{w} \cup \vec{z})$. Since the sub-formulas added are
  tautologies, the matrix defined at this point remains unchanged, and so
  $\lambda (\vec{x})$ and $\theta (\vec{x})$ are equivalent. The idea is that
  $\op{tag}$ adds a tower of conjunctions of tautologies each referencing
  exactly one variable from $\vec{w} \cup \vec{z}$, and so encoding the order of
  the sequence of variables $\vec{w} \cup \vec{z}$ in the syntax of the formula.
  When we define a circuit from this formula, each input to the rank gate will
  be `tagged' with a unique assignment to $\vec{w} \cup \vec{z}$, preventing two
  inputs from being syntactically-equivalent, and so ensuring the circuit is
  transparent.

  % We also relabel the bound variables in $\lambda(\vec{x})$ such that there
  % are
  % no two sub-formulas $\psi$ and $\psi'$ such that $\psi$ and $\psi'$ are
  % equal
  % up to a renaming of those variables that are bound in $\psi$ or $\psi'$.

  % Let $\psi (\vec{y}) \in \FOrk$ and $\alpha \in [n]^{\vec{y}}$. Let
  % $\psi[\alpha]$ be the result of substituting $\vec{y}$ for $\alpha(y)$ in
  % $\psi(\vec{y})$. If $\psi$ is of the form $y_1 = y_2$. Then $\psi [\alpha] =
  % 1$ if $\alpha(y_1) = \alpha (y_2)$ and $\psi[\alpha] = 0$ otherwise.

  We note that $\lambda(\vec{x})$ has the same width as $\theta(\vec{x})$, and
  that the number of sub-formulas in $\lambda(\vec{x})$ is at most $\vert
  \cl{\theta} \vert (5 + 4\width(\theta))$. We now construct a circuit $C$ and
  show that $C$ translates $\lambda(\vec{x})$ (and hence $\theta(\vec{x})$) at
  $n$.

  % We first
  % pre-proccess the formula $\theta(\vec{x})$

  % Let $\op{no-op}(y) := (R(y, y) \land (\neg R(y, y)))$. Let $\vec{y} := (y_1,
  % \ldots, y_k)$ be a vector of variables and $\op{add-tag} (\vec{y}) :=
  % \op{no-op}(y_1) \land (\op{no-op}(y_2) \land ( \op{no-op}(y_2) \land (
  % \ldots
  % \land (\op{no-op}(y_k)) \ldots)))$. From $\theta (\vec{x})$ we construct a
  % new
  % formula $\theta' (\vec{x})$ by replacing each sub-formula $\phi(\vec{y})$ of
  % the form $\rank^r_p \vec{w} \vec{z} . \psi$ by the sub-formula $\rank^r_p
  % \vec{w}\vec{z} . ((\forall u . u = u) \land \phi) \land \op{add-tag}(\vec{w}
  % \cup \vec{z})$.

  Let $\vec{y}$ be a sequence of variables. We abuse notation and write $\alpha
  :\vec{y} \rightarrow A$ to indicate that $\alpha$ maps the variables in the
  sequence $\vec{y}$ to elements of the set $A$.

  Fix $n \in \nats$. Let $\psi \in \FOrk$ and $\alpha \in [n]^{\free {\psi}}$.
  We let $\psi[\alpha]$ be the result of substituting each occurrence of the
  free variable $y \in \free{\psi}$ in $\psi$ with $\alpha(y)$. If $\psi$ is of
  the form $y_1 = y_2$. Then $\psi [\alpha] = 1$ if $\alpha(y_1) = \alpha (y_2)$
  and $\psi[\alpha] = 0$ otherwise. We call $\psi[\alpha]$ a \emph{ground
    formula}. For each $\psi \in \cl{\theta}$ let $G_\psi := \{g_{\phi[\alpha]}
  : \alpha \in [n]^{\free{\psi}} \}$. Let $G := \bigcup_{\psi \in \cl{\lambda}}
  G_{\psi}$. We define $\Sigma$, $\Lambda$ and $L$ as follows. Let $\Sigma (g_1)
  = 1$ and $\Sigma (g_0) = 0$. Let $g := g_{\psi[\alpha]} \in G$.

  \begin{itemizens}
  \item If $\psi[\alpha] = T(\vec{a})$ for some $T \in \tau$ and $\vec{a} \in
    \arty(T)$, then $\Sigma (g) = T$ and $\Lambda_T (g) = \vec{a}$.
  \item If $\psi[\alpha]$ is $0$ or $1$, then $\Sigma (g) = \psi[\alpha]$.
  \item If $\psi$ is of the form $\rank^r_p \vec{w} \vec{z}. \phi$, let $A =
    [n]^{\vert \vec{w}\vert}$ and $B = [n]^{\vert \vec{z}\vert}$, and $\Sigma
    (g_{\psi, \alpha}) = \rank[A, B]$. Let $L(g) : A \times B \rightarrow G$ be
    defined by $L(g)(\vec{a},\vec{b}) = g_{\phi[\beta]}$, where $\beta \in
    [n]^{\free{\phi}}$, and where $\alpha$, and the assignments
    $\alpha^{\vec{w}}_{\vec{a}}$ and $\alpha^{\vec{z}}_{\vec{b}}$, are
    compatible with $\beta$.
  \item If $\psi$ is of the form $Q z . \phi(\vec{y}, z)$, for $Q \in \{\forall,
    \exists\}$, then if $Q = \forall$ we have $\Sigma (g) = \land [n]$,
    otherwise $\Sigma (g) = \lor [n]$. Let $L(g) : [n] \rightarrow G$ be defined
    for $i \in [n]$ by $L(g)(i) = g_{\phi[\beta_i]}$, where $\beta_i \in
    [n]^{\free{\psi}}$, and where both $\alpha$ and $\alpha^{z}_{i}$ are
    compatible with $\beta_i$.
  \item If $\psi$ is of the form $\phi_1 \land \phi_2$, then $\Sigma(g) =
    \land[2]$ and $L(g) : [2] \rightarrow G$ is defined for $i \in [2]$ by
    $L(g)(i) = g_{\phi_i[\beta_i]}$, where $\beta_i \in [n]^{\free{\psi}}$ and
    $\beta_i$ is compatible with $\alpha$. The same approach is used in the
    disjunctive case.
  \item if $\psi$ is of the form $\neg \phi$ then $\Sigma (g) = \neg$ and $L(g):
    [1] \rightarrow G$, where $L(g)(1) = g_{\phi[\alpha]}$.
  \end{itemizens}

  We let $\Omega : [n]^k \rightarrow G$ be defined by $\Omega (\vec{a}) =
  g_{\lambda [\alpha]}$, where $\alpha := \alpha^{\vec{x}}_{\vec{a}}$, for all
  $\vec{a} \in [n]^k$. Suppose $\Omega (\vec{a}) = g_c$, where $ c \in {0,1}$.
  Then $\lambda$ (and $\theta$) is of the form $(y_1 = y_2)$. But then the
  vocabulary of $\theta$ is empty, contradicting our assumption. Suppose $\Omega
  (\vec{a}) = \Omega(\vec{b})$ for $\vec{a}, \vec{b} \in [n]^k$. Then
  $g_{\lambda[\alpha^{\vec{x}}_{\vec{a}}]} =
  g_{\lambda[\alpha^{\vec{x}}_{\vec{b}}]}$, and so $\alpha^{\vec{x}}_{\vec{a}} =
  \alpha^{\vec{x}}_{\vec{b}}$, and $\vec{a} = \vec{b}$. It follows that $\Omega$
  is injective. Let $T \in \tau$, and let $g_{\psi[\alpha]}, g_{\phi[\beta]} \in
  G$ be such that $\Lambda_T (g_{\psi[\alpha]}) = \Lambda_T (g_{\phi[\beta]})$.
  Then $\psi[\alpha] = \Lambda_T (g_{\psi[\alpha]}) = \Lambda_T
  (g_{\phi[\beta]}) = \phi[\beta]$, and so $g_{\psi[\alpha]} = g_{\phi[\beta]}$.
  It follows that $\Lambda_T$ is injective. We thus have that $C := \langle G,
  \Omega, \Sigma, \Lambda, L \rangle$ is a circuit of order $n$.

\begin{claim}
  The circuit $C$ is symmetric.
\end{claim}
\begin{proof}
  Let $\pi_\sigma : G \rightarrow G$ be a function that fixes constant gates for
  such that for all non-constant gates $g_{\psi[\alpha]} \in G$, $\pi_{\sigma}
  g_{\psi [\alpha]} = g_{\psi[\sigma \alpha]}$. It is easy to see that
  $\pi_\sigma$ is a bijection. We now show that $\pi_{\sigma}$ is an
  automorphism of the circuit extending $\sigma$. Let $g_{\psi[\alpha]} := g \in
  G$. Since the formula is unchanged by $\pi_\sigma$, we have $\Sigma(g) =
  \Sigma (\pi_{\sigma} g)$. Suppose $g$ is an input gate. If $g$ is a constant
  gate then $\pi_\sigma g = g$. If $g$ is a relational gate, then $\psi$ is of
  the form $T(\vec{y})$ for some relation symbol $T$ and $\Lambda_T
  (g_{\psi[\alpha]}) = \alpha \vec{y}$, and $\Lambda (\pi_{\sigma}
  g_{\psi[\alpha]}) = \Lambda g_{\psi[\sigma \alpha]} = \sigma \alpha \vec{y} =
  \sigma \Lambda_T (g_{\psi[\alpha]})$. Suppose $g$ is an internal gate. From
  the construction of the circuit we have that if $g_{\phi[\beta]} \in H_g$ then
  $\phi$ must be an immediate sub-formula of $\psi$ and so $g_{\phi[\sigma
    \beta]}$ is a child of $g_{\psi[\sigma \alpha]} = \pi_{\sigma} g$. It
  follows that $H_{\pi_\sigma g} = \pi_\sigma H_g$. If $g$ is a symmetric gate,
  this is sufficient to conclude that $\pi_{\sigma} L(g)$ is
  isomorphism-equivalent to $L(\pi_{\sigma} g)$. If $g$ not a symmetric gate
  then it is a rank gate, and so $\psi$ is of the form $\rank^r_p \vec{w}
  \vec{z} \phi$. Let $\vec{a} \in A$ and $\vec{b} \in B$, and note that
  $\pi_\sigma L(g) (\vec{a}, \vec{b}) = \pi_{\sigma} g_{\phi[\beta]}$, where
  $\beta \in [n]^{\free {\phi}}$ is compatible with $\alpha$,
  $\alpha^{\vec{w}}_{\vec{a}}$ and $\alpha^{\vec{z}}_{\vec{b}}$. It follows that
  $\sigma \beta$ is compatible with $\sigma \alpha$, $\sigma
  \alpha^{\vec{w}}_{\vec{a}} = \alpha^{\vec{w}}_{\sigma \vec{a}}$ and
  $\alpha^{\vec{z}}_{\vec{b}} = \alpha^{\vec{z}}_{\sigma \vec{b}}$. We thus have
  that $L(\pi_{\sigma} g)(\sigma \vec{a}, \sigma \vec{b}) = g_{\phi[\sigma
    \beta]} = \pi_{\sigma} g_{\phi[\beta]} = \pi_{\sigma} L(g)(\vec{a},
  \vec{b})$. Suppose $g$ is an output gate, and let $\vec{a} \in [n]^k$ and
  $\alpha := \alpha^{\vec{x}}_{\vec{a}}$. We have that $\pi_{\sigma} \Omega
  (\vec{a}) = \pi_\sigma g_{\lambda[\alpha]} = g_{\lambda[\sigma \alpha]} =
  \Omega (\sigma \alpha \vec{x}) = \Omega (\sigma \vec{a})$. This concludes the
  proof of the claim.
\end{proof}

Let $\mathcal{A}$ be a $\tau$-structure with universe $U$ of size $n$ and let
$\gamma \in [n]^{\underline{U}}$. It can be shown that for all $\psi \in
\cl{\lambda}$ and $\alpha \in U^{\free{\psi}}$, $\mathcal{A} \models
\psi[\alpha]$ if, and only if, $C[\gamma \mathcal{A}](g_{\psi[\gamma\alpha]}) =
1$. It follows that $\vec{a} \in U^k$ is an element of $Q(\theta(\vec{x}))$ if,
and only if, $\mathcal{A} \models \lambda[\alpha^{\vec{x}}_{\vec{a}}]$ if, and
only if, $C[\gamma \mathcal{A}](g_{\lambda[\alpha^{\vec{x}}_{\gamma \vec{a}}]})
= 1$. Since $g_{\lambda [\alpha^{\vec{x}}_{\gamma \vec{a}}]} = \Omega (\gamma
\vec{a})$, we have that $C$ computes the query defined by $\theta(\vec{x})$ on
structures of size $n$.

\begin{claim}
  The circuit $C$ is transparent.
\end{claim}
\begin{proof}
  If $C$ has symmetric gates then $C$ is transparent. Suppose $C$ contains a
  non-symmetric gate. Let $g_{\phi[\alpha]}:= g \in G$ be a rank gate and
  suppose $\psi$ is of the form $\rank^r_p \vec{w} \vec{z}. \zeta$, where $\zeta
  = \zeta_1 \land \zeta_2$ and where, $\zeta_1 = ((\forall u. u = u) \land
  \phi)$, and $\zeta_2 = \land (\op{tag}(\vec{w} \cup \vec{z}))$. Let $\beta_1 ,
  \beta_2 \in [n]^{\free{\zeta}}$ both compatible with $\alpha$, and let $h_1 =
  g_{\zeta[\beta_1]}$ and $h_2 = g_{\zeta[\beta_2]}$. For $i \in [2]$ we have
  that $H_{h_i} = \{g_{\zeta_1[\beta_i]}, g_{\zeta_2[\beta_i]}\}$. Since
  $\zeta_1$ contains a sub-formula of the form $(u = u)$, while $\zeta_2$
  contains no sub-formula with equality, we have that a child of a child of
  $g_{\zeta_1[\beta_1]}$ is a constant gate while none of the children of the
  children of $g_{\zeta_2[\beta_2]}$ are constant gates, and thus $g_{\zeta_1
    [\beta_1]} \not\equiv g_{\zeta_2[\beta_2]}$. It follows that $h_1 \equiv
  h_2$ if, and only if, both $g_{\zeta_1 [\beta_1]} \equiv g_{\zeta_1
    [\beta_2]}$ and $g_{\zeta_2[\beta_1]} \equiv g_{\zeta_2[\beta_2]}$.
  Moreover, the fact that $\zeta_2$ contains an ordered sequence of conjunctions
  of tautologies, with each tautology referencing a different variable,
  guarantees that $g_{\zeta_2[\beta_1]} \equiv g_{\zeta_2[\beta_2]}$ if, and
  only if, $\beta_1$ and $\beta_2$ agree on $\free{\zeta_2}$. But we already
  have that $\beta_1$ and $\beta_2$ are compatible with $\alpha$, and so agree
  on $\free{\zeta_1}$. It follows that $g_{\zeta_2[\beta_1]} \equiv
  g_{\zeta_2[\beta_2]}$ if, and only if, $\beta_1 = \beta_2$. This allows us to
  conclude that $g_{\zeta[\beta_1]} \equiv g_{\zeta[\beta_2]}$ if, and only if,
  $\beta_1 = \beta_2$ if, and only if, $g_{\zeta[\beta_1]} =
  g_{\zeta[\beta_2]}$. It follows that $L(g)$ is injective, and no two $h_1, h_2
  \in H_g$ are syntactically-equivalent. We conclude that $g$ has unique labels,
  and so $C$ is transparent.
\end{proof}

We notice that $\vert C \vert \leq n^{\width(\lambda)} \vert \cl{\lambda} \vert
\leq n^{\width(\theta)} \vert \cl{\theta} \vert (5 + \width(\theta))$. It is not
hard to see that $C$ (and $\lambda(\vec{x})$) can be constructed in a number of
steps bounded by a polynomial in the size of $C$. This completes the proof of
the result.

% = C[\gamma \mathcal{A}](\Omega(\gamma \vec{a}))$

% Since $h_{}$Since $\pi_\sigma$ is a bijection, we have that $f$ is a
% bijection. Moreover, we have that $\pi L(g) (a) = L(g) $

% We also have that $\mathcal{A} \models \psi[\gamma \alpha]$ if, and only if,
% $T^{\mathcal{A}} (\gamma \alpha \vec{y})$ if, and only if, $C[\gamma
% \mathcal{A}](g) = 1$

% Suppose $g$ is an internal gate, and we have both the results are true for all
% gates that come before $g$ in the topological ordering of the circuit. Suppose
% $\psi$ is an application of is an application of a logical connective ($land$
% or $\lor$) or negation ($\neg$). We then have that $g$ is a symmetric gate,
% and $f : H_{g} \rightarrow H_{\pi_\sigma g}$ defined by $f(h) = \pi_\sigma h$
% is a bijection.

% $\pi_\sigma g_{\psi[\alpha]} = \Lambda_T ()$, for some formula $\psi$
% containing the relation symbol $T$ and assignment $\alpha$, where $\alpha$
% assigns the free variables in $\psi$ to $\vec{a}$. Then $\pi_\sigma g =
% \Lambda^{-1}_T (\sigma \Lambda_T (g)) = g_{\psi[\sigma \alpha]}$. Suppose $g$
% is an internal gate. If $g = g_{psi[\alpha]}$ is an application of a logical
% connective ( $land$ or $\lor$) or negation ($\neg$) then $\pi_{\sigma}g =
% g_{\psi[\sigma \alpha]}$. If $g$

% It can be shown by induction on the structure of the circuit that all
% $g_{\psi[\alpha]} \in G$

% We should note that $C'$ is not, in general, a circuit. This is because it has
% (potentially) more than three constant gates, and perhaps more than one
% relational gate for each assignment and relation symbol. Let $C$ be formed
% from $C'$ by identifying all gates $g$ and $g'$ such that (i) $g$ and $g'$ are
% constant gates and $\Sigma' (g) = \Sigma'(g')$, or (ii) $g$ and $g'$ are
% relational gates, $T := \Sigma' (g) = \Sigma' (g') $, and $\Lambda_T' (g) =
% \Lambda_T'(g')$. We now have $C := \langle G, \Omega, \Sigma, \Lambda, L
% \rangle$. If the domain of $\Omega'$ includes a constant gate, then
% $\theta(\vec{x}) = \Lambda (\vec{x}) = (x_1 = x_2)$

% Notice that, formally, $C'$ is not a circuit, in that it contains too many
% relational gates.

% We now show that $C$ translates $\theta(\vec{x})$ at $n$. It is obvious that
% $C$ has order $n$. For $\sigma \in \sym_n$ let $\pi_\sigma : G \rightarrow G$
% be a defined such that $\pi_{\sigma} (g) = g$ if $g \in G_c$ and
% $\pi_\sigma(g_{\psi, \alpha}) = g_{\psi, \sigma \alpha}$ for all $\psi \in
% \cl{\lambda}$ and $\alpha \in n^{\free{\psi}}$. It can be shown that
% $\pi_\sigma$ is an automorphism of $C$ extending $\sigma$. It follows that $C$
% is symmetric.

% Let $\mathcal{A}$ be a $\tau$-structure with a universe $U$ of size $n$ and
% let $\gamma U^{\underline{[n]}}$. Let $\psi (\vec{y}) \in \cl{\lambda}$ and
% let $\alpha \in [n]^{\vec{y}}$. It can be shown by induction on the structure
% of the circuit that for all $\psi (\vec{y})$ and $\alpha \in [n]^{\vec{y}}$,
% $C[\gamma \mathcal{A}](g_{\psi, \alpha}) = 1$ if, and only if, $\mathcal{A}
% \models \psi [\gamma (\alpha \vec{y})]$. It follows that

% Let $\vec{a} \in [n]^k$, and $\alpha \in [n]^{\vec{x}}$ be defined such that
% $\alpha (x_i) = a_i$ for all $i \in [k]$. Then $C[\Omega]$

% It is obvious that $C$ has order $n$. Let $\sigma \in \sym_n$. Let $\pi_\sigma
% : G \rightarrow G$ be a defined such that $\pi_{\sigma} (g) = g$ if $g \in
% G_c$ and $\pi_\sigma(g_{\psi, \alpha}) = g_{\psi, \sigma \alpha}$ for all
% $\psi \in \cl{\lambda}$ and $\alpha \in n^{\free{\psi}}$. It can be shown that
% $\pi_\sigma$ is an automorphism of $C$ extending $\sigma$. It follows that $C$
% is symmetric.

% Moreover, let $\alpha \in [n]^{\free{\theta}}$ be an assignment that maps the
% sequence of variables $\vec{x}$ to the tuple $\vec{a} \in [n]^k$. It follows
% that $\Omega (\vec{a}) = g_{\lambda, \alpha}$ evaluates to $1$ If $C$ does not
% contain a rank gate then it is transparent. Suppose $C$ does contain a rank
% gate. Let $g := g_{\psi, \alpha} \in G$ be a rank gate and suppose $\psi
% (\vec{y}) := \rank^r_p \vec{w} \vec{z}. \zeta$, where $\zeta = \zeta_1 \land
% \zeta_2$ and $\zeta_1 = ((\forall u. u = u) \land \phi)$ and $\zeta_2 = \land
% (\op{tag}(\vec{w} \cup \vec{z}))$. Suppose $h_1, h_2 \in H_g$ are
% syntactically-equivalent. It follows that there exist assignments $\beta_1,
% \beta_2 \in [n]^{\vec{w} \cup \vec{z} \cup {y}}$ such that $\alpha \subseteq
% \beta_1$ and $\alpha \subseteq \beta_2$ and $h_1 = g_{\zeta, \beta_1}$ and
% $h_2 = g_{\zeta, \beta_2}$. Then $H_{h_i} = \{g_{\zeta_1, \beta_i},
% g_{\zeta_2, \beta_i}\}$ for $i \in [2]$. Note that a child of a child of
% $g_{\zeta_1, \beta_1}$ is a constant gate while none of the children of the
% children of $g_{\zeta_2, \beta_2}$ are constant gates, and thus $g_{\zeta_1,
% \beta_1} \not\equiv g_{\zeta_2, \beta_2}$ are not syntactically-equivalent. It
% follows that $h_1 \equiv h_2$ if, and only if, both $g_{\zeta_1, \beta_1}
% \equiv g_{\zeta_1, \beta_2}$ and $g_{\zeta_2, \beta_1} \equiv g_{\zeta_2,
% \beta_2}$. However, the construction of $\zeta_2$ (and the fact that $\beta_1$
% and $\beta_2$ agree on the variables in $\vec{y}$) guarantees that
% $g_{\zeta_2, \beta_1} \equiv g_{\zeta_2, \beta_2}$ if, and only if, $\beta_1 =
% \beta_2$ are equal. It follows that $h_1 = g_{\zeta, \beta_1} = g_{\zeta,
% \beta_2} = h_2$. The construction of the circuit guarantees that $g$ has
% injective labels. It follows that $g$ has unique labels, and the circuit $C$
% is transparent.

% and $h := g_{\phi, \beta}$ and $h' := g_{\phi, \beta'}$ are
% syntactically-equivalent. From the construction of $\theta'$ we have that
% $\phi$ must be of a particular form. Suppose $\phi = \eta \land
% \op{add-tag}(\vec{w} \cup \vec{z})$. Let $\gamma$ and $\gamma'$ be the
% restrictions of $\beta$ and $\beta'$ to $\vec{w} \cup \vec{z}$, respectively,
% and let $\zeta(\vec{w}\cup\vec{z}) := \op{add-tag}(\vec{w} \cup \vec{z})$. The
% $\op{add-tag} (\vec{w} \cup \vec{z})$

% We notice that $\vert C \vert \leq n^{\width(\lambda)} \vert \cl{\lambda}
% \vert \leq n^{\width(\theta)} \vert \cl{theta} \vert (5 + \width(\theta))$. It
% can be shown that $C$ (and $\lambda(\vec{x})$ can be constructed in a number
% of steps bounded by a polynomial in the size of $C$.

% If $C'$ contains no rank gates then output $C := C'$. Suppose $C'$ contains at
% least one rank-gate. We note add to $C'$ a gadget that ensures that no two
% input gates to a rank gate can be syntactically-equivalent. Let $\psi \in
% \cl{\theta}$, and suppose $\psi (\vec{y}) = \rank^r_p \vec{w} \vec{z}. \phi$.
% We may think of the assignments to the variables $\vec{w}$ and $\vec{z}$ as
% being the row and column indexes of the matrix. We

% Then for $\alpha \in [n]^{\free (\psi)}$, $H_{g_{\psi, \alpha}}$ consists of
% gates of the form $g_{\phi, \beta}$, where $\beta \in [n]^{\vec{w} \cup
% \vec{z} \cup \vec{y}}$ and $\alpha subseteq \beta$.

% Let $T_{\psi, \alpha} : H_{g_{\psi, \alpha}} \rightarrow [n]^{\vert \vec{w}
% \vert + \vert \vec{z} \vert}$ be defined by $T_{\psi, \alpha} (g_{\phi,
% \beta}) = \beta (\vec{z} \cup \vec{w})$ for all $g_{\phi, \beta} \in H_g$.
% What $T_{\psi, \alpha}$ does is assign to each child of $g_{\psi, \alpha}$ an
% ordered sequence formed Let $T$ be the union of all $T_{\psi}$ and $H$ the
% union of all $G_{\psi}$ for all $\psi \in \cl{\theta}$ an application of the
% rank operator. Let $C := \op{tag-all}(C', R, H, T)$. We note that
% $\op{tag-all}$ may be computed in time polynomial in the size of the input
% circuit.

% Let $a \in [n]$ and let $\op{no-op} (a) := R(a, a) \lor (\neg (R(a, a))$. For
% $\psi (\vec{y})$

% % Let $(b_1, \ldots, b_m) := \beta (\vec{w} \cup \vec{z})$. The $\op{tag-all}
% % function replaces each input $g_{\phi, \beta}$ of the rank gate $g_{\psi,
% % \alpha}$ with the circuit $g_{\phi, \beta} \land \op{no-op}(b_1) \land
% % (\op{no-op}(b_2) \land ( \op{no-op}(b_2) \land (\ldots \land
% % (\op{no-op}(b_m))
% % \ldots)))$. This construction does not effect the evaluation of any rank
% % gate
% % in the cirucit and ensures that no two to $g_{\psi,\alpha}$ in $C$ are
% % syntactically-equivilent.

% $of the form $$R(a, a) \lor (\neg R(a, a))$ that encode the assignment to
% $\vec{w}\cup \vec{z}$.

% We have that $C$ is transparent, has order $n$, and $C$ and $C'$ decide the
% same query. It remains to show that $C$ is still symmetric.

% However, the $\op{tag-all}$ operation adds a tower of tautologies of the form
% $R(a, \ldots, a) \lor (\neg R (a, \ldots, a))$

% Let $g := g_{\psi, \alpha} \in G$ be a rank gate and let $h, h' \in H_g$.
% Suppose $\psi (\vec{y}) := \rank^r_p \vec{w} \vec{z}. \phi$ and $h := g_{\phi,
% \beta}$ and $h' := g_{\phi, \beta'}$ are syntactically-equivalent. From the
% construction of $\theta'$ we have that $\phi$ must be of a particular form.
% Suppose $\phi = \eta \land \op{add-tag}(\vec{w} \cup \vec{z})$. Let $\gamma$
% and $\gamma'$ be the restrictions of $\beta$ and $\beta'$ to $\vec{w} \cup
% \vec{z}$, respectively, and let $\zeta(\vec{w}\cup\vec{z}) :=
% \op{add-tag}(\vec{w} \cup \vec{z})$. The $\op{add-tag} (\vec{w} \cup \vec{z})$

% The structure of this formula gives us that $h$ is syntactically-equivalent to
% $h'$ if, and only if, $g_{\zeta , \beta'}$

% We also note that $C$ has order $n$.
\end{proof}
